### 2018-07-06 paper/5-implementations.tex
Simplify the description of suspending

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Shake build system; stores verifying traces
shake :: (Ord k, Hashable v) => Build Monad (VT k v) k v
shake = suspending vtRebuilder
\end{minted}
\vspace{0mm}
\begin{minted}[fontsize=\small]{haskell}
-- A task rebuilder based on verifying traces
vtRebuilder :: (Eq k, Hashable v) => Rebuilder Monad (VT k v) k v
vtRebuilder key value task = Task $ \fetch -> do
    upToDate <- verifyVT key (hash value) (fmap hash . fetch) =<< get
    if upToDate then return value else do
        (newValue, deps) <- track task fetch
        modify $ recordVT key (hash newValue) [ (k, hash v) | (k, v) <- deps ]
        return newValue
\end{minted}
\vspace{0mm}
\begin{minted}[fontsize=\small]{haskell}
-- A suspending task scheduler
suspending :: Ord k => Scheduler Monad i i k v
suspending rebuilder tasks target store = fst $ execState (fetch target) (store, Set.empty)
  where
    fetch :: k -> State (Store i k v, Set k) v
    fetch key = do
        done <- gets snd
        case tasks key of
            Just task | key `Set.notMember` done -> do
                value <- gets (getValue key . fst)
                let newTask :: Task (MonadState i) k v
                    newTask = rebuilder key value task
                newValue <- liftRun newTask fetch
                modify $ \(s, d) -> (updateValue key value newValue s, Set.insert key d)
                return newValue
            _ -> gets (getValue key . fst) -- fetch the existing value
\end{minted}
\vspace{0mm}
\begin{minted}[fontsize=\small]{haskell}
-- Run a Task using Store to obtain the i
liftRun :: Task (MonadState i) k v
        -> (k -> State (Store i k v, Set k) v) -> State (Store i k v, Set k) v
\end{minted}
\vspace{-3mm}
\caption{An implementation of \Shake using our framework.}\label{fig-shake-implementation}
\vspace{-6mm}
\end{figure}

@@

The \hs{suspending} scheduler uses a recursive function \hs{build} that builds
a given \hs{key}, making sure not to duplicate work when called on the same
\hs{key} again in future. To achieve that, it keeps track of keys that have
already been built in a set \hs{done}~\hs{::}~\hs{Set}~\hs{k}.
% , running the computation in the
% \hs{State}~\hs{(@@Store}~\hs{i}~\hs{k}~\hs{v,}~\hs{Set}~\hs{k)} monad.

Given a non-input \hs{key} that has not yet been built, we use the supplied
\hs{rebuilder} to embed the build information~\hs{i} into the \hs{task}.
We then execute the obtained \hs{newTask} with the \hs{fetch} callback, which
performs the following three steps when given a dependency \hs{k}:

\begin{itemize}
    \item Build \hs{k} by recursively calling the \hs{build} function. Note
          that if \hs{k} is dirty, this will effectively suspend the build of
          the current \hs{key} until \hs{k} is brought up to date.
    \item Save any new traces that have been generated when building \hs{k}.
    \item Read the (up to date) value of \hs{k} from the store and return it.
\end{itemize}

These steps are run in the
\hs{State}~\hs{(@@Store}~\hs{i}~\hs{k}~\hs{v,}~\hs{Set}~\hs{k)}
monad. To make \hs{MonadState}~\hs{i} access the \hs{i} inside the \hs{Store}
we use the helper function \hs{liftRun} (which uses a \hs{newtype} to provide a \hs{MonadState}
instance that sees through into the \hs{Store}).
The \hs{newValue} obtained by running the \hs{newTask} is
stored, and the \hs{key} is added to the set \hs{done}.

\subsection{Cloud build systems: \Bazel, \CloudBuild, \Cloud \Shake, \Buck and \Nix}\label{sec-implementation-cloud}

>>>>

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Shake build system; stores verifying traces
shake :: (Ord k, Hashable v) => Build Monad (VT k v) k v
shake = suspending vtRebuilder
\end{minted}
\vspace{0mm}
\begin{minted}[fontsize=\small]{haskell}
-- A task rebuilder based on verifying traces
vtRebuilder :: (Eq k, Hashable v) => Rebuilder Monad (VT k v) k v
vtRebuilder key value task = Task $ \fetch -> do
    upToDate <- verifyVT key (hash value) (fmap hash . fetch) =<< get
    if upToDate then return value else do
        (newValue, deps) <- track task fetch
        modify $ recordVT key (hash newValue) [ (k, hash v) | (k, v) <- deps ]
        return newValue
\end{minted}
\vspace{0mm}
\begin{minted}[fontsize=\small]{haskell}
-- A suspending task scheduler
suspending :: Ord k => Scheduler Monad i i k v
suspending rebuilder tasks target store = fst $ execState@\,@(fetch target)@\,@(store,@\,@Set.empty)
  where
    fetch :: k -> State (Store i k v, Set k) v
    fetch key = do
        done <- gets snd
        case tasks key of
            Just task | key `Set.notMember` done -> do
                value <- gets (getValue key . fst)
                let newTask :: Task (MonadState i) k v
                    newTask = rebuilder key value task
                newValue <- liftRun newTask fetch
                modify $ \(s, d) -> (putValue key newValue s, Set.insert key d)
                return newValue
            _ -> gets (getValue key . fst) -- fetch the existing value
\end{minted}
\vspace{0mm}
\begin{minted}[fontsize=\small]{haskell}
-- Run a task using a callback that operates on a larger state (implementation omitted)
liftRun :: Task (MonadState i) k v
        -> (k -> State (Store i k v, Set k) v) -> State (Store i k v, Set k) v
\end{minted}
\vspace{-3mm}
\caption{An implementation of \Shake using our framework.}\label{fig-shake-implementation}
\vspace{-6mm}
\end{figure}

@@

The \hs{suspending} scheduler uses a recursive \hs{fetch} callback, defined
similarly to the \hs{busy} build system (\S\ref{sec-general-build}), that builds
a given \hs{key}, making sure not to duplicate work when called on the same
\hs{key} again in future. To achieve that, it keeps track of keys that have
already been built in a set \hs{done}~\hs{::}~\hs{Set}~\hs{k}. Given a non-input
\hs{key} that has not yet been built, we use the supplied \hs{rebuilder} to
embed the build information~\hs{i} into the \hs{task}. We then execute the
obtained \hs{newTask} by passing it the \hs{fetch} function as a callback for
building dependencies: the \hs{newTask} will therefore be suspended while its
dependencies are being brought up to date. The \hs{newValue} obtained by running
the \hs{newTask} is stored, and the \hs{key} is added to the set \hs{done}.

The \hs{fetch} computation runs in the
\hs{State}~\hs{(@@Store}~\hs{i}~\hs{k}~\hs{v,}~\hs{Set}~\hs{k)} monad. To make
\hs{MonadState}~\hs{i} access the \hs{i} inside the \hs{Store} we use the helper
function \hs{liftRun} (which uses a \hs{newtype} to provide a \hs{MonadState}
instance that sees through into the \hs{Store}).

\subsection{Cloud build systems: \Bazel, \CloudBuild, \Cloud \Shake, \Buck and \Nix}
\label{sec-implementation-cloud}
